---
title: "Homework 4 - Report"
output: html_document
date: "2023-02-10"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(foreign)
library(DescTools)


```

```{r include=FALSE}
BD <- read.spss("BaseCovid_trabalho4.sav", to.data.frame = TRUE)
```

#### A group of researchers wants to know how to characterize the world countries with more cases and deaths by COVID-19, per million inhabitants. To analyze the data inherent to this study, consider the SPSS database BaseCovid_trabalho4.sav.
The research team decided to analyze the association between countries' median age being above or below 30 years and their total number of COVID-19 cases per million population.

##### Do these data suggest that there is a relationship between the median age of countries and their number of cases of COVID-19 per million population?

Let's make a crosstab of our data:
```{r echo=FALSE}
cross_age_cases <- xtabs(~BD$total_cases_per_million_10_tercil+BD$median_age_2cat)
cross_age_cases
```

To check for a relationship between the median age of countries and their number of cases of COVID-19 per million population we want to perform a chi-square test. 
To do that, we must first test the assumptions:
```{r echo=FALSE}
expected <- ExpFreq(cross_age_cases)
```

We can verify that all expected values are >5, so we can proceed with our chi-square test of independence.
```{r echo=FALSE}
qui <- chisq.test(BD$total_cases_per_million_10_tercil, BD$median_age_2cat)
qui
```
With a p-value of 3.299e-07, we reject the null hypothesis that these variables are independent and, as such, there is a statistically significant relationship between them.

