---
title: "Homework 1 - Report"
output: html_document
date: "2023-02-10"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse)
library(ggplot2)
library(readxl)
library(freqtables)
library(naniar)
library(stats)
library(moments)

```

```{r include=FALSE}
asma <- read_xlsx("Asma_GDH.xlsx")
```

### Validate the database, looking for potential missing values, impossible values, outliers, etc.

First, let's do a simple table with summary statistics:
```{r echo=FALSE}
miss_var_summary(asma) #Returns a tibble of missing values count and percent by variable
```
We can see that there are no missing values in our data.
Let's look for impossible values and outliers:
```{r echo=FALSE}

asma %>%
  summarise(n = n(), 
            mean = mean(`Idade (anos)`, na.rm = TRUE), 
            sd = sd(`Idade (anos)`, na.rm = TRUE),
            stderr = sd/sqrt(n), 
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(`Idade (anos)`, na.rm = TRUE),
            min=min(`Idade (anos)`, na.rm = TRUE), 
            max=max(`Idade (anos)`, na.rm = TRUE),
            IQR=IQR(`Idade (anos)`, na.rm = TRUE),
            kurt= kurtosis(`Idade (anos)`),
            skew = skewness(`Idade (anos)`))

#quantitativa variables: idade, dias internamento,num procedimentos
age_box_plot <- boxplot(asma$`Idade (anos)`,
                        ylab="Age (years)",
                        main="Boxplot of Partcipant Ages",
                        col="orange",
                        border="brown") #No outliers or impossible values

```

We can see that our data has no outliers or impossible values regarding particpant's age.  
The same procedure would be applied to any other quantitative variables we would want to explore.


### Transform the variable "Idade (anos)" in a binary variable 
(0 -- Abaixo da Mediana; 1 -- Acima da Mediana)

```{r eval=FALSE}
median_age <- median(asma$`Idade (anos)`)

asma <- asma %>% 
  mutate(age_bin = ifelse(`Idade (anos)`>median_age,1,0))

```

### Identify your analysis objectives at the univariate level. Characterize the sample globally, generating the tables and graphs you find most relevant.

```{r echo=FALSE}
freq_NUTS <- freq_table(asma, `NUTS III`)
freq_NUTS

bar_NUTS <- ggplot() + 
  geom_bar(asma, mapping=aes(`NUTS III`, fill=`NUTS III`))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
bar_NUTS

asma %>%
  summarise(n = n(), 
            mean = mean(`Numero de Procedimentos`, na.rm = TRUE), 
            sd = sd(`Numero de Procedimentos`, na.rm = TRUE),
            stderr = sd/sqrt(n), 
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(`Numero de Procedimentos`, na.rm = TRUE),
            min=min(`Numero de Procedimentos`, na.rm = TRUE), 
            max=max(`Numero de Procedimentos`, na.rm = TRUE),
            IQR=IQR(`Numero de Procedimentos`, na.rm = TRUE),
            kurt= kurtosis(`Numero de Procedimentos`),
            skew = skewness(`Numero de Procedimentos`))

hist_procedure <- hist(asma$`Numero de Procedimentos`, breaks=50, right=F,
                       main="Histogram of number of procedures",
                       xlab="Number of procedures",
                       col="#d2d67a")

```

### Now identify your analysis objectives at the bivariate level. Characterize the sample globally, generating the tables and graphs you find most relevant. Perform at least one of each of the following bivariate variable crossings:

-   Qualitative variable x Qualitative variable

```{r echo=FALSE}
#Quali x Quali --> Districts X Episode Type
cross_dist_epi <- xtabs(~asma$Distrito + asma$`Tipo Episodio`)
cross_dist_epi
```
-   Qualitative variable x Quantitative variable
Here we are analyzing the type of asthma diagnose and the age of the patient:
```{r echo=FALSE}
#Quanti x Quali
# Type of asthma dx and age
summary_asth_age <- asma %>% 
  group_by(`Tipo diagnostico Asma`) %>% 
  summarize(mean = mean(`Idade (anos)`),
            median = median(`Idade (anos)`),
            min = min(`Idade (anos)`),
            max = max(`Idade (anos)`),
            stdev = sd(`Idade (anos)`),
            var = stdev^2,
            kurt= kurtosis(`Idade (anos)`),
            skew = skewness(`Idade (anos)`),
            interquart = IQR(`Idade (anos)`))
  
summary_asth_age
```
-   Quantitative variable x Quantitative variable
Here we are analyzing the number of procedures and hospitalization days
```{r echo=FALSE}
#Quanti x Quanti
# No of procedures and intensive care days
plot(asma$`Numero de Procedimentos`, asma$`Dias de internamento`,
     main="Graphical representation of the correlation \nbetween hospitalization days and number of procedures",
     xlab="Number of procedures",
     ylab="Hospitalization days",
     col="#141d82")
correlation <- cor.test(asma$`Numero de Procedimentos`, asma$`Dias de internamento`)
correlation
```
  We find that there is a statistically significant correlation between the number of procedures and hospitalization days, with a coefficient of 0.508.