---
title: "Homework 3 - Report"
output: word_document
date: "2023-02-10"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(tidyverse)
library(ggplot2)
library(foreign)
library(qqplotr)
library(car)
library(emmeans)
library(DescTools)
library(FSA)
library(PMCMRplus)

```
```{r echo=FALSE}
BD <- read.spss("TPC_3_Caracterizacao_Concelhos_Saude.sav", to.data.frame = TRUE)
```

### We intend to compare the North and Lisbon regions regarding the number of physicians. Please indicate, justifying, what inferential statistical methodologies would be appropriate for this purpose? Name the hypothesis tests used, comment on and interpret the results. 

```{r echo=FALSE}
NL_BD <- BD %>% 
  base::subset(NUTS_II == "Norte" | NUTS_II == "Área Metropolitana de Lisboa")

#Produce boxplots and visually check for outliers
ggplot(NL_BD, aes(x = NUTS_II, y = Médicos, fill = NUTS_II)) +
  stat_boxplot(geom ="errorbar", width = 0.5) +
  geom_boxplot(fill = "light blue") + 
  stat_summary(fun.y=mean, geom="point", shape=10, size=3.5, color="black") + 
  ggtitle("Boxplots of Lisbon and North Region Groups") + 
  theme_bw() + theme(legend.position="none")

#Produce descriptive statistics by group
NL_BD %>% select(Médicos, NUTS_II) %>% group_by(NUTS_II) %>% 
  summarise(n = n(), 
            mean = mean(Médicos, na.rm = TRUE), 
            sd = sd(Médicos, na.rm = TRUE),
            stderr = sd/sqrt(n), 
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(Médicos, na.rm = TRUE),
            min=min(Médicos, na.rm = TRUE), 
            max=max(Médicos, na.rm = TRUE),
            IQR=IQR(Médicos, na.rm = TRUE))

# QQ plots
ggplot(data = NL_BD, mapping = aes(sample = Médicos, color = NUTS_II, fill = NUTS_II)) +
  stat_qq_band(alpha=0.5, conf=0.95, qtype=1, bandType = "ts") +
  stat_qq_line(identity=TRUE) +
  stat_qq_point(col="black") +
  facet_wrap(~ NUTS_II, scales = "free") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") + theme_bw() 
```
First, we need to test for assumptions:
Normal distribuition in each group, and homogeneity of variances

Starting with Shapiro-Wilk to test for normality:
```{r echo=FALSE}

NL_BD %>%
  group_by(NUTS_II) %>%
  summarise(`W Statistic` = shapiro.test(Médicos)$statistic,
            `p-value` = shapiro.test(Médicos)$p.value)

```
Both p-values are <0.05, meaning we reject the null hypothesis, so this data is not normally distributed
We then need to do a non-parametric test: Mann-Whitney U 

Before that, and for the sake of practice, lets test for homogeneity of variances
```{r echo=FALSE}

lev1<-leveneTest(Médicos ~ NUTS_II, data=NL_BD)
print(lev1)


```

we get a p-value of 0.01281, meaning we reject the null hypothesis, so between group variances are unequal

```{r echo=FALSE}
# moving on to the Mann whitney U test
mann_whitney <- wilcox.test(Médicos ~ NUTS_II, data=NL_BD, na.rm=TRUE, paired=FALSE, exact=FALSE, conf.int=TRUE)
mann_whitney
```
We get a p-value of 3.405e-06, meaning we reject the null hypothesis that distributions are equal and conclude that there is a significant difference in the number of doctors of both regions.

### A researcher argues that there are statistically significant differences between the average values per county for the number of physicians and the number of nurses in Portugal. Do you agree with the researcher's statement? Identify the appropriate statistical methodologies to apply, comment and interpret the results.

```{r echo=FALSE}

BD <- BD %>% 
  mutate(diff = Médicos - Enfermeiros)

BD %>% select(Médicos)  %>% 
  summarise(n = n(),
            mean = mean(Médicos, na.rm = TRUE), 
            sd = sd(Médicos, na.rm = TRUE),
            stderr = sd/sqrt(n), 
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(Médicos, na.rm = TRUE),
            min=min(Médicos, na.rm = TRUE), 
            max=max(Médicos, na.rm = TRUE),
            IQR=IQR(Médicos, na.rm = TRUE))

BD %>% select(Enfermeiros) %>% 
  summarise(n = n(),
            mean = mean(Enfermeiros, na.rm = TRUE), 
            sd = sd(Enfermeiros, na.rm = TRUE),
            stderr = sd/sqrt(n), 
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(Enfermeiros, na.rm = TRUE),
            min=min(Enfermeiros, na.rm = TRUE), 
            max=max(Enfermeiros, na.rm = TRUE),
            IQR=IQR(Enfermeiros, na.rm = TRUE))

# QQ plot of diff
ggplot(data = BD, mapping = aes(sample = diff)) +
  stat_qq_band(alpha=0.25, conf=0.95, qtype=1, bandType = "boot", B=5000, fill="red") +
  stat_qq_line(identity=TRUE) +
  stat_qq_point(col="black") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") + theme_bw() 

# Boxplot of diff
ggplot(BD, aes(x = "", y = diff)) +
  stat_boxplot(geom ="errorbar", width = 0.5) +
  geom_boxplot(fill = "light blue") + 
  stat_summary(fun.y=mean, geom="point", shape=10, size=3.5, color="black") + 
  ggtitle("Boxplot of doctors - nurses differences") + 
  theme_bw() + theme(legend.position="none")
```

```{r echo=FALSE}
# Testing for normality in diff
shapiro.test(BD$diff)
```
With a p-value of <2.2e-16, we reject the null hypothesis and conclude that the difference between groups is not normally distributed.

Since we rejected normality, we must perform a non parametric test: Wilcoxon test
```{r echo=FALSE}
#  
wilcox <- wilcox.test(BD$Médicos, BD$Enfermeiros,na.rm=TRUE, paired=TRUE, exact=FALSE, conf.int=TRUE)
wilcox #p-value <2.2e-16, we reject the null hypothesis
```

With a p-value of <2.2e-16, we reject the null hypothesis and conclude that there is a statistically significant difference between the number of physicians and the number of nurses.

### You want to compare all regions regarding the number of physicians. Please indicate, justifying, which inferential statistical methodologies would be appropriate for this purpose? Name the hypothesis tests used, comment and interpret the results.


```{r echo=FALSE}
Med_BD <- BD %>% 
  select(Médicos, NUTS_II)

ggplot(Med_BD, aes(x = NUTS_II, y = Médicos, fill = NUTS_II)) +
  stat_boxplot(geom ="errorbar", width = 0.5) +
  geom_boxplot(fill = "light blue") + 
  stat_summary(fun.y=mean, geom="point", shape=10, size=3.5, color="black") + 
  ggtitle("Boxplots of Lisbon and North Region Groups") + 
  theme_bw() + theme(legend.position="none")

#Produce descriptive statistics by group
Med_BD %>% select(Médicos, NUTS_II) %>% group_by(NUTS_II) %>% 
  summarise(n = n(), 
            mean = mean(Médicos, na.rm = TRUE), 
            sd = sd(Médicos, na.rm = TRUE),
            stderr = sd/sqrt(n), 
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(Médicos, na.rm = TRUE),
            min=min(Médicos, na.rm = TRUE), 
            max=max(Médicos, na.rm = TRUE),
            IQR=IQR(Médicos, na.rm = TRUE))

# QQ plots
ggplot(data = Med_BD, mapping = aes(sample = Médicos, color = NUTS_II, fill = NUTS_II)) +
  stat_qq_band(alpha=0.5, conf=0.95, qtype=1, bandType = "ts") +
  stat_qq_line(identity=TRUE) +
  stat_qq_point(col="black") +
  facet_wrap(~ NUTS_II, scales = "free") +
  labs(x = "Theoretical Quantiles", y = "Sample Quantiles") + theme_bw() 
```
Since we want to compare multiple groups, we want to perform an ANOVA test or Kruskall-Wallis
First, we need to test for assumptions:
# Normal distribuition in each group; 
# Homogeneity of variances

Starting with Shapiro-Wilk to test for normality
```{r echo=FALSE}

Med_BD %>%
  group_by(NUTS_II) %>%
  summarise(`W Statistic` = shapiro.test(Médicos)$statistic,
            `p-value` = shapiro.test(Médicos)$p.value)
# 
```

All of them present a p-value <0.05,we reject the null hypothesis, and conclude that this data is not normally distributed.
We then need to do a non-parametric test: Kruskal-Wallis, but before that, and for the sake of practice, lets test for homogeneity of variances

```{r echo=FALSE}
lev2<-leveneTest(Médicos ~ NUTS_II, data=Med_BD)
print(lev2)

```
We get a p-value of 0.0001606, meaning we reject the null hypothesis, so between group variances are unequal.

Let's finally move onto the test we want:
```{r echo=FALSE}
m1<-kruskal.test(Médicos ~ NUTS_II, data=Med_BD)
print(m1) 
```
With a p-value of 1.19e-11,we reject the null hypothesis, and conclude that there are differences in median values between groups

We must then move onto post-hoc analysis:
```{r echo=FALSE}
#Dunn's Kruskal-Wallis post-hoc test
posthocs1<-dunnTest(Médicos ~ NUTS_II, data=Med_BD, method="holm")
print(posthocs1)

#Dwass, Steel, Critchlow, Fligner post-hoc test
posthocs2<-dscfAllPairsTest(Médicos ~ NUTS_II, data=Med_BD)
print(posthocs2)
```

